Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-18 talk/pepm2011/paper.tex
more small tweaks, still missing a more aggressive rewrite

\subsection{Running Example}

@@

\begin{figure}
\begin{lstlisting}[mathescape]
# arguments to the trace: $p_{0}$, $p_{1}$
# inside f: res.add(y)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = int_add($i_{2}$, $i_{3}$)
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)

@@

The main insight to improve the code shown in the last section is that objects
in category 1 do not survive very long -- they are used only inside the loop and
there is no other outside reference to them. The idea for improving
the code is to analyze which objects fall in category 1 and therefore do
not have to be allocated at all.

@@

In the example from last section, the following operations in the upper half
of Figure~\ref{fig:unopt-trace} produce two
static objects, and can be completely removed from the optimized trace:

@@

In this section we want to give a formal description of the semantics of the
traces and of the optimizer and liken the optimization to partial evaluation.
We focus on the operations for manipulating dynamically allocated objects,
as those are the only ones that are actually optimized. Without loss of
generality we also consider only objects with two fields in this section.

@@

The static heap is a partial function from $V^*$ into the
set of static objects, which are triples of a type and two elements of $V^*$.
A variable $v^*$ is in the domain of the static heap $S$ as long as the
optimizer does not need to become dynamic XXX. The object $S(v^*)$ describes
what is statically known about the object, \ie its type and its fields. The
fields of objects in the static heap are also elements of $V^*$ (or null, for
short periods of time).

>>>>

\subsection{Running Example}
\label{sub:example}

@@

\begin{figure}
\lstset{
    numbers = right,
    numberstyle = \tiny,
    numbersep = 0pt
}
\begin{lstlisting}[mathescape]
# arguments to the trace: $p_{0}$, $p_{1}$
# inside f: res.add(y)
guard_class($p_{1}$, BoxedInteger)
    # inside BoxedInteger.add
    $i_{2}$ = get($p_{1}$, intval)
    guard_class($p_{0}$, BoxedInteger)
        # inside BoxedInteger.add__int
        $i_{3}$ = get($p_{0}$, intval)
        $i_{4}$ = int_add($i_{2}$, $i_{3}$)
        $p_{5}$ = new(BoxedInteger)
            # inside BoxedInteger.__init__
            set($p_{5}$, intval, $i_{4}$)

@@

The main insight to improve the code shown in the last section is that objects
in category 1 do not survive very long -- they are used only inside the loop
and there is no other outside reference to them. Therefore the optimizer
identifies objects in category 1 and removes the allocation of these objects,
and all operations manipulating them.

@@

In the example from Section~\ref{sub:example}, the following operations in the upper half
of Figure~\ref{fig:unopt-trace} produce two
static objects, and can be completely removed from the optimized trace:

@@

In this section we want to give a formal description of the semantics of the
traces and of the optimizer and liken the optimization to partial evaluation.
We focus on the operations for manipulating dynamically allocated objects,
as those are the only ones that are actually optimized. We also consider only
objects with two fields in this section, generalizing to arbitrary many fields
is straightforward.

@@

The static heap is a partial function from $V^*$ into the
set of static objects, which are triples of a type and two elements of $V^*$.
The object referenced by a variable $v^*$ is static, if $v^*$ is in the domain
of the static heap $S$. The object $S(v^*)$ describes
what is statically known about the object, \ie its type and its fields. The
fields of objects in the static heap are also elements of $V^*$ (or null, for
short periods of time).
